---
title: "Lab 3"
#author: "Zoe Drajem"
date: "Math 241, Week 3"
output:
  pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
# Do not modify this chunk.
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


```

```{r}
libs <- c('tidyverse','knitr','viridis', 'mosaic','mosaicData','babynames', 'Lahman','nycflights13','rnoaa')
for(l in libs){
  if(!require(l,character.only = TRUE, quietly = TRUE)){
    message( sprintf('Did not have the required package << %s >> installed. Downloading now ... ',l))
    install.packages(l) 
  }
  library(l, character.only = TRUE, quietly = TRUE)
}

```



## Due: Friday, February 16th at 8:30am

## Goals of this lab

1. Practice creating functions.
1. Practice refactoring your code to make it better!  Therefore for each problem, make sure to test your functions.


### Problem 1: Subset that R Object

Here are the R objects we will use in this problem (`dats`, `pdxTreesSmall` and `ht`).    

```{r}
library(tidyverse)
library(pdxTrees)
library(mosaicData)

pdxTrees <- get_pdxTrees_parks()
# Creating the objects
dats <- list(pdxTrees  = head(pdxTrees),
             Births2015 = head(Births2015),
             HELPrct = head(HELPrct), 
             sets = c("pdxTrees", "Births2015", 
                      "HELPrct"))

pdxTreesSmall  <- head(pdxTrees)

ht <- head(pdxTrees$Tree_Height, n = 15)

```

a. What are the classes of `dats`, `pdxTreesSmall` and `ht`?

b. Find the 10th, 11th, and 12th values of `ht`
```{r}
ht[10:12]

```


c. Provide the `Species` column of `pdxTrees` as a data frame with one column.

```{r}
pdxTrees_species <- pdxTrees %>%
  select(Species)
```


d. Provide the `Species` column of `pdxTrees` as a character vector.

```{r}
vector <- pdxTrees$Species 
```


e. Provide code that gives us the second entry in `sets` from `dats`.

```{r}
dats$sets[2]

```


f. Subset `pdxTreesSmall` to only `Douglas-fir` and then provide the `DBH` and `Condition` of the 4th `Douglas-fir` in the dataset.  (Feel free to mix in some `tidyverse` code if you would like to.)

```{r}
pdxTreeNew <- pdxTreesSmall %>%
  filter(Common_Name == "Douglas-Fir") 

pdxTreeNew$DBH[4]
pdxTreeNew$Condition[4]

```




### Problem 2: Function Creation

Figure out what the following code does and then turn it into a function. For your new function, do the following:

* Test it.
* Provide default values (when appropriate).
* Use clear names for the function and arguments.
* Make sure to appropriately handle missingness.
* Generalize it by allowing the user to specify a confidence level.
* Check the inputs and stop the function if the user provides inappropriate values.


```{r}
library(pdxTrees)

create_confidence_interval <- function(x, alpha = 0.05){
  if (!is.numeric(x)){
    stop('Unfortunately this function only works for numeric input.\n',
         'You have provided an object of class:', class(x)) 
  }
  
  if (!is.numeric(alpha)){
    stop('Unfortunately this function only works for numeric input.\n',
         'You have provided an object of class:', class(x))
  }
  
  x<- x[!is.na(x)]
  
  length <- length(x)
  mean <- mean(x)
  se <- sd(x)/sqrt(length)
  t_star <- qt(p = 1-(alpha/2), df = length - 1)
  lower_interval <- mean - t_star*se
  upper_interval <- mean + t_star*se
  
  return(c(lower_interval,upper_interval))
}

create_confidence_interval(pdxTrees$DBH, alpha = 0.01)


```





### Problem 3: Wrapper Function for your `ggplot`


While we (i.e. Math 241 students) all love the grammar of graphics, not everyone else does. So for this problem, we are going to practice creating wrapper functions for `ggplot2`.

Here's an example of a wrapper for a histogram.  Notice that I can't just list the variable name as an argument.  The issue has to do with how many of the `tidyverse` functions evaluate the arguments. Therefore we have to quote (`enquo()`) and then unquote (`!!`) the arguments.  (If you want to learn more, go [here](https://tidyeval.tidyverse.org/).)

```{r}
# Minimal viable product working code
ggplot(data = pdxTrees, mapping = aes(x = DBH)) +
  geom_histogram()

# Shorthand histogram function
histo <- function(data, x){
  x <- enquo(x)
  ggplot(data = data, mapping = aes(x = !!x)) +
    geom_histogram()
}

# Test it
histo(pdxTrees, DBH)
```

a. Edit `histo()` so that the user can set 

* The number of bins
* The fill color for the bars
* The color outlining the bars
```{r}
histo <- function(data, x, color = "pink", fill = "pink", bins = 30){
  x <- enquo(x)
  color <- enquo(color)
  fill <- enquo(fill)
  ggplot(data = data, mapping = aes(x = !!x, color = !!color, fill = !!fill)) +
    geom_histogram(bins = bins) 
}

histo(pdxTrees, DBH)
```



b. Write code to create a basic scatterplot with `ggplot2`. Then write and test a function to create a basic scatterplot.

```{r}
scat <- function(data, x, y){
  x <- enquo(x)
  y <- enquo(y)
  ggplot(data = data, mapping = aes(x = !!x, y = !!y)) +
    geom_point() 
}

scat(pdxTrees, Latitude, Longitude)
```


c. Modify your scatterplot function to allow the user to ...

* Color the points by another variable.
* Set the transparency.

```{r}
scatter <- function(data, x, y, color, alpha = 0.1) {
  x <- enquo(x)
  y <- enquo(y)
  color <- enquo(color)
  ggplot(data = data, mapping = aes(x = !!x, y = !!y, color = !!color)) +
    geom_point(alpha = alpha) 
}

scatter(pdxTrees, Latitude, Longitude, Condition, alpha = 0.01)
```


d. Write and test a function for your favorite `ggplot2` graph.

```{r}
box <- function(data, x, y){
  x <- enquo(x)
  y <- enquo(y)

  ggplot(data = data, mapping = aes(x = !!x, y = !!y)) +
    geom_boxplot()
}

box(pdxTrees, DBH, Condition)

```



### Problem 4: Functioning `dplyr`


a. Take the following code and turn it into an R function to create a **conditional proportions** table.  Similar to `ggplot2`, you will need to quote and unquote the variable names.  Make sure to test your function!

```{r}
conditional_prop_table <- function(data, x, y){
  x <- enquo(x)
  y <- enquo(y)

  data %>%
    count(!!x, !!y) %>%
    group_by(!!x) %>%
    mutate(prop = n/sum(n)) %>%
    ungroup()
}

conditional_prop_table(pdxTrees, Native, Condition)
```


b.  Write a function to compute the mean, median, sd, min, max, sample size, and number of missing values of a quantitative variable by the categories of another variable.  Make sure the output is a data frame (or tibble).  Don't forget to test your function.

```{r}
summary <- function(data, x, y){
  x <- enquo(x)
  y <- enquo(y)

  data %>%
    group_by(!!y) %>%
    summarise(
      mean = mean(!!x),
      median = median(!!x),
      sd = sd(!!x), 
      min = min(!!x),
      max = max(!!x), 
      sample_size = n(), 
      n_missing = sum(is.na(!!x))
    )
}

summary(pdxTrees, DBH, Condition)
```



### Problem 5: another `babynames` exercise

Write a function called grab_name that, when given a **name** ***and a year*** as an argument, returns the rows from the `babynames` data frame in the `babynames` package that match that name for that year (and returns an error if that name and year combination does not match any rows). Run the function once with the arguments **Ezekiel and 1883** and once with **Ezekiel and 1983**.

```{r, eval=TRUE}

grab_name <- function(myname, myyear) {
  if (is.element(myname, babynames$name)) { babynames %>%
      filter(name == myname) %>%
       filter(year == myyear) %>%
      return()
    }
  else {
    stop("Name not found")
  }
}

grab_name("Ezekiel", 1883)

grab_name("Ezekiel", 1983)
```